Bayesian exponential random graph modeling of KWL Burial network

### 2020-10-16 analysis/paper/paper.Rmd
reorganized paras

<!--archaeology case-->
The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and accumulation of them in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society. Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups. In addition to prestige, European colonial powers could have provided military support to Indigenous allies. For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals. In this study, we will use this model of competition to explain the changes in social inequality at Kiwulan, an Iron Age site in northesatern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century.

# Exponential random graph models

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that an observed network is randomly generated depending on the presence or absence of other edges. ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008]. In statistical models, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011]. For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby. Similarly, for network dependencies, reciprocity represents the likelihood between two actor's to be tied because of shared properties, and transitivity means two actors tend to be linked via a relationship with a third actor in common. We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. Network specification under assumptions and contexts is useful for exploring archaeological network. For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs. Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively. However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data. In addition, the complex computational processes, such as intractable normalizing constant and degeneracy due to extreme graphs, limits the widely use of ERGMs. Here, we used ERGMs in a Bayesian framework that allows for parameter inference using Markov chain Monte Carlo (MCMC) algorithm which avoid the need for computationally intensive calculations of the normalizing constants [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006].

## Bayesian modeling

<!--Bayesian inference for Exponential Random Graph Models--> 
Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019]. This provides better estimations for complex social network models with heterogeneous data [@Caimo2017]. For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced by using Bayesian modelling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010].
- weighted data

Using the R software, we apply Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan. We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group. We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. We ask the questions: 1. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks. 3. How the relations among burials may affect the quantity of burial goods? We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

# Archaeological background

@@

<!--archaeological setting-->
We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (1350-1850 AD) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007]. The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated. The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison. Those burials are oriented in an east-west direction on the north side of the house structures indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement.

Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials. However, there is a debate over whether the differentiation between burials hints an economic inequality supporting an egalitarian society [@Hsieh2012], or political inequality indicating a stratified society [@Cheng2008]. @Cheng2008 interprets the unequal distribution of prestige materials, especially the gold-foil beads between burials, as evidence for hierarchy. However, @Hsieh2012 suggests a relatively egalitarian society based on the comparative analysis of frequency and proportion of all burial goods. She found that the burials with rich goods were usually associated with elders, which might indicate achieved, rather than inherited, status. The earliest record of the use of prestige goods in burials can be traced back around the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006]. For example, the Spanish document by a Spanish Priest described that the Spanish soldier used agate beads to exchange materials from local Indigenous people since they were valuable in their culture [@LiandWu2006]. In addition, there was a custom that people would bury imported materials with deceased to display their status and influence [@LiandWu2006]. Despite we observed the economic differentiation in burial goods and the high value of imported goods according to historical documents, the degree of inequality over time, and which variable decides the changes remain unclear. Analyzing and comparing burial data from different phases using network analysis approaches will give an insight into the degree and changes in social inequality. This is important to understand the social reactions of Kiwulan residents in the context of the foreign contact.

@@

We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. In our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, presence of burial rituals, and overall burial goods quantity index. Burial ritual is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

Our hypothesis is that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power. This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with more social inequality. If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by those social processes. Network analysis provides a way to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

## Model specification 
<!--Network construction-->

We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis. ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In ERGMs, ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics. By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks. Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007]. This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.

Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. We test our hypothesis by measuring reciprocity, transitivity, popularity, and homophily structure effects. We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence. This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality. We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any.

## Bayesian modelling

```{r ergm-terms-table}
ergm_terms <- readxl::read_excel(here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
kableExtra::kable(ergm_terms,
             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
```

<!--Bayesian approaches-->

We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008]. Although exponential random graph models can describe a network by capturing the dependency structure, it is difficult to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian]. Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. To overcome these limitations in ERGM, Bayesian model applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011]. This alleviate above computational problems and gives a better convergence results. By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters. Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian]. The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.

<!--Model fitting and comparison-->

Markov chain Monte Carlo (MCMC) algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].

<!-- Goodness-of-fit assessment-->

>>>>

<!--archaeology case-->
The archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and accumulation of them in residential area around the same time as European contact in the 17th century [@Chen2007], which might indicate the pursuit of prestige or wealth in the society. Those prestige goods include imported ornaments and porcelains that entered northeastern Taiwan through regional exchange networks involved the Europeans, Han Chinese traders, and Indigenous groups. In addition to prestige, European colonial powers could have provided military support to Indigenous allies. For example, @Voss2005's study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local Indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values of imported goods, might lead to increased social inquqality due to competition among individuals. In this study, we use this pericolonial model to explain the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the presence of the Europeans, including the Spanish and the Dutch, in the 17th century and the Chinese in the 19th century. Our purpose is to identify changes in network structure of burials before and after the foreign contacts, and test our hypothesis that increased inequality observed in social network after the European arrival. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

We applied Bayesian SNA to burial data from Kiwulan, an Iron Age site in northeast Taiwan. We view individual burials as actors or nodes in the social network with ties drawn between burials when the presence of the same variables that can represent the same "community" or social group. We test hypotheses about the distribution of influence in the network to identify network or corporate mode conditions. We ask the questions: 1. Do the observed burial data indicate a more clustered network than a distribution of random networks with similar qualities? 2. Whether European colonial activities in 17 century Taiwan resulted in the emergence of social inequality in an Indigenous society can be detected by burials networks. 3. How the relations among burials may affect the quantity of burial goods? We expect the network to present a higher degree of clustering (e.g. the number of completed triangles in the network) in the European contact period that might hint at social heterogeneity during this period. This study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. A Bayesian approach enables the efficient quantification for uncertainty, parametrization, and model evaluation of social network metrics.

# Exponential random graph models in a Bayseisan framework

<!--current approaches to network analysis-->
Exponential Random Graph Models (ERGMs) are an important family of statistical models for networks that allows directly modeling for the formation of edges, or ties, between nodes [@Robins2007]. The assumption is that possible ties in a network are random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. ERGMs differs from traditional network approaches that edges are only formed according to their corresponding predictors separately without taken into account the interactions among edges [@Morris2008]. In ERGMs, networks are viewed as dependent variables, the formation of a tie is also influenced by network dependencies, and the attributes of nodes/edges [@Snijders2011]. For example, nodes with similar attributes is more likely to form a relationship, such as friendship between people with the same hobby. Ties form a small structure in a network called a graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with a similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations represent the structure or the property of a network and can be expressed by network statistics. We can model those network statistics as direct functions of ties, where the probability of occurrence of ties can be specified and hypothesized [@Morris2008; @Robins2007]. By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks. Such distribution consists of a large number of possible networks that provides a statistical inference for an observed network [@Robins2007].  This helps to understand whether an observed network shows significantly more or less property of interest than the random networks we modeled.

ERGMs can describe a network by capturing the dependency structure by specifying parameters under assumptions and contexts, which is useful for exploring archaeological network. For instance, @Amati2019 demonstrated how to reconstruct of three different networks consisting of 15 sites (AD 100 to 400) in the Caribbean to explore their interaction mechanisms using ERGMs. Three network models test different hypotheses for the relationship between sites based on proximity, inter-cultural, and exchange network to explain the presence of hub sites they observed. They found that the presence of hubs can be efficiently explained by multiple interdependent mechanisms instead of only one model exclusively. However, they also pointed out some limitations of ERGMs, such as static outcome and sensitivity to missing data. In addition, it is difficult in ERGMs to estimate model parameters and interpret the result due to the intractable normalizing constant and model degeneracy [@Caimo2014; @Jin2013]. A normalizing constant is an important step to make probability distribution to integrate to one, which becomes harder to compute with larger set of networks [@Caimo2017bayesian]. Degeneracy is another issue that probability models tend to overestimate a small number of extreme graphs by assigning too much weight, such as empty (all nodes unconnected) or complete graphs (all nodes connected) [@Schweinberger2011; @Caimo2014]. Those complex computational processes in ERGMs can be also resolved by implementing ERGMs in a Bayesian framework.

<!--Bayesian inference for Exponential Random Graph Models--> 
Bayesian approach to ERGMs is an efficient computational tool for social network analysis because it incorporates prior information about the network effects into the model and offers uncertainty quantification by evaluating the posterior distribution of the parameters associated with the network effects [@Caimo2015; @Nemmers2019]. This provides better estimations for complex social network models with heterogeneous data [@Caimo2017]. For archaeological data, missing data could be a problem leading to misinterpretation, which can be also reduced with Bayesian modeling that provides average 80% correct prediction for the ties and allow 25% false with a third of data missing [@Koskinen2010]. To overcome these limitations in ERGMs, Bayesian modeling applies Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm to estimate the double intractable posteriors [@Caimo2011]. MCMC algorithms can produce approximate maximum likelihood estimators and graph statistics for further specification of transitivity and heterogeneity, this enables ERGMs to deal with complicated dependence patterns with a ease [@Snijders2006]. This alleviate above computational problems and gives a better convergence results. By fitting an ERMG with MCMC algorithm, a Bayesian approach provides posterior probability that incorporates our sample data and prior information through summary statistics from estimates of the ERGM parameters. Posterior probability enables the estimation the effect of ERGM parameter by looking at the posterior mean and 95% credible intervals, which can replace p-value for accessing a null hypothesis [@Caimo2017bayesian]. The criteria is that the odds ratio greater than one means positive effect of ERGM parameter, while odds ratio less than one represents negative effect.

# Kiwulan in northeastern Taiwan

@@

<!--archaeological setting-->
We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (1350-1850 AD) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007]. The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated. The burials are mostly located in the middle section of the excavation area, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison. Those burials are oriented in an east-west direction on the north side of the residential area indicated by post-holes and `in-situ` wooden posts, indicating a well organized space arrangement. Previous studies point out a presence of uneven distribution of prestige goods across burials without an agreement that whether this uneven distribution hints vertical social differences. @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society. However, @Hsieh2012 suggested a relatively egalitarian based on the comparative analysis of frequency and proportion of all burial goods. She found that the burials with high valued burial goods were usually associated with elders, which might indicate achieved, rather than inherited, status.

The earliest record of the use of prestige goods in burials can be traced back to the time when the European arrival in the 17th century according to historical records from the Spanish [@Borao2009; @LiandWu2006]. For example, the document written by a Spanish Priest described that the Spanish soldier used carnelian beads to exchange natural resources with local Indigenous people since their high values in Indigenous culture [@LiandWu2006]. In addition, there was a custom that people would bury imported beads and ceramics with deceased to display their status and influence [@LiandWu2006]. Despite we observed the uneven distribution of prestige goods in burials, the degree of differentiation across burials over time remain unclear. The discord over the discussion of Kiwulan social organization could be associated with chronological differences that has not yet been well identified. That is, the observed uneven distribution may be explained as result of social change from less social differences to more social differences when Indigenous societies involved in the complex trade network stimulated by the Europeans. We assume that the changes would be supported by chronological differences in the structure of burial networks. Analyzing and comparing networks between different phases relative to cultural contact will give an insight into the degree and changes in social differences. This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.

@@

We analyzed burial data collected from published excavation reports, and original fieldwork notes from Kiwulan. We do not include the burials of the Chinese phase in our network analysis due to small sample size. We assume that the pericolonial impact hypothesis would be supported by changes in network structures build upon flow of resources, prestige goods here. In our case, burials are nodes that are linked when they have goods in common. Given that gold-foil beads and agate beads commonly found across burials with substantial differences in counts, we assigned those prestige goods into three levels, high, medium, and low, to indicate quantity based on the distribution of their counts. That means if burial 1 and burial 2 both have high quantity of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we assign a tie between burials when they possess each type of goods in common. Node attributes here include osteological data and mortuary practices, such as age, gender, ritual practice, and overall burial goods quantity index. Ritual practice is identified by locally made ceramics placed above graves that suggests funeral feasting according to historical records. Quantity index is an important attribute for economical inequality that we assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the total number of burial goods. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

Our pericolonial impact hypothesis is that restricted grave goods or accumulation of goods present in a few individuals after the European presence in the 17th century, because of increased social inequality at Kiwulan due to differential access high value goods. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to the interest of burial goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals for building their personal status and power. If the social inequality gradually increased in Kiwulan, then we would observe a network with higher degree (popularity), less transitivity (cohesion), and strong inequality based on age difference (achieved status). Network analysis provides a way to examine the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships through chronological comparison.

## Model specification in a Bayesian framework
```{r ergm-terms-table}
ergm_terms <- readxl::read_excel(here("analysis", "data", "raw_data", "ergm_term.xlsx")) 
kableExtra::kable(ergm_terms,
             caption = "The parameters of exponential random graph models for a undirected network to corresponding with archaeological evidence.")
```

<!--Network construction and specification-->
We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis of social change using the R programming language with bergm package [@Caimo2014], which built upon ergm package for simulation and model specification [@Hunter2008; @Morris2008; @Handcock2008]. Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with archaeological interpretations. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. Based on our hypothesis, we model a network with increased social inequality that represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects. We modeled several node attributes for homophily effects, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried close to each other. Our model infers the presence of a few individuals as centers having more relations with others, which hints the emergence of social inequality. We expect that observed network after the European arrival will have a better match with our model compared with the network before European contact.

<!---Based on our hypothesis, we model a network using parameters including reciprocity, transitivity, popularity, and homophily structure effects. We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence. This structure infers the presence of a few individuals as centers having more relations with others with a decrease in small and multiple groups, which hints the emergence of social inequality. We also consider homophily effects based on node attributes, including age, gender, ritual, and the degree of wealth, to explore the correlation between the tendency of connected nodes and the network ties overall. This can imply the major effect that contributes to the network that we observed and give an insight into the underlying mechanism of social inequality if any. --->

<!--Model fitting and prior setting--> 
After the specification, the network is simulated in a Baysesian framework using Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithm is an estimation of posterior distribution of interest through directly random sampling the posterior without assuming the prior is a normal distribution. We can obtain a distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on algorithms. This enables a random or stochastic simulation in a long run without depending on the previous move. More chains ensures a more desirable distribution close to the target distribution that we study, which is called the convergence of the Markov chains. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and than switch to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000.

<!-- Goodness-of-fit assessment-->
The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017].

<!--Since our original data implies higher transitivity due to potentially multiple groups having the same kind of burial good in common, we specified the prior based on our observation with high transitivity and low density. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical approach to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network. Our observed networks, pre-European and post-European networks, followed the normal distribution that is a suitable prior model in network analysis that assumes networks with low density and high transitivity commonly observed in real world [@Caimo2017] -->
